Slitasje og kjørespor

### 2022-11-28 DRAFT_slitasjeOgKjørespor.Rmd
update paths

<!-- Load all you dependencies here -->
```{r setup, include=FALSE}
library(knitr)
library(sf)
knitr::opts_chunk$set(echo = TRUE)
```

@@

#### Nature type mapping
This indicator uses the dataset `Naturtyper etter Miljødirektoratets Instruks`, which can be found [here](https://kartkatalog.geonorge.no/metadata/naturtyper-miljoedirektoratets-instruks/eb48dd19-03da-41e1-afd9-7ebc3079265c).
The dataset has already been [cleaned](#naturtype1) and [initial analyses on thematic representativity](#naturtype2) has been done.

```{r, eval=T}
dat <- readRDS("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/Mdir_Naturtyper_cleanedALK.rds")
```

#### Vegetation zones (bioregions)

#### Regions
<!-- In case you need to map the indicator value to regions, you can do that here. Remove this chapter if not relevant. -->
Importing a shape file with the regional delineation. 
```{r}
#reg <- sf::st_read("data/regions.shp")
#st_crs(reg)
```

>>>>

<!-- Load all you dependencies here -->
```{r setup, include=FALSE}
library(knitr)
library(sf)
library(tmap)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```

@@

#### Nature type mapping
This indicator uses the dataset `Naturtyper etter Miljødirektoratets Instruks`, which can be found [here](https://kartkatalog.geonorge.no/metadata/naturtyper-miljoedirektoratets-instruks/eb48dd19-03da-41e1-afd9-7ebc3079265c).
See [here](#naturtype) for a detailed description of tha dataset.

```{r, eval=T}

dir <- substr(getwd(), 1,2)

path <- ifelse(dir == "C:", 
               "P:/41201785_okologisk_tilstand_2022_2023/data/Natur_Naturtyper_NiN_norge_med_svalbard_25833.gdb",
               "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/Natur_Naturtyper_NiN_norge_med_svalbard_25833.gdb")

dat <- sf::st_read(dsn = path)
```

We also have a separate summary file where the nature types are manually mapped to the correct NiN-main types. We can use this to find the nature types associated with the NiN main types that we are interested in.
```{r}
nt <- readRDS("data/naturetypes/natureType_summary.rds")
```

We are only interested in a few NiN main types and their mapping units (see [here](https://ninanor.github.io/ecosystemCondition/naturtype.html) for justification)
```{r}
main <- c("T12",
          "T18",
          "T21",
          "T31",
          "T33",
          "T34",
          "V1",
          "V3")
```

```{r}
nt2 <- nt[nt$NiN_mainType %in% main,]
```

Sub-setting the main data file:
```{r}
dat2 <- dat[dat$naturtype %in% nt2$Nature_type,]
```

Fix duplicated name for an ecosystem
```{r}
dat2$hovedøkosystem[dat2$hovedøkosystem=="Naturlig åpne områder i lavlandet"] <- "Naturlig åpne områder under skoggrensa"
```

```{r, fig.cap="An overview of the naturetypes for which we will calculate the indicator. Colours refer to the main ecosystem affiliation."}
ggplot(data = dat2, aes(x = naturtype, fill = hovedøkosystem))+
  geom_bar()+
  coord_flip()+
  theme_bw(base_size = 12)+
  guides(fill = "none")+
  xlab("")+
  ylab("Number of localities")
```

# GRUK

This variable is also recorded in GRUK

#### Vegetation zones (bioregions)
I want to look at if the indicator values can be averages over smaller and more meaningfu areas (ecological homogeneous areas) than the regions.
```{r}

path <- ifelse(dir == "C:", 
               "R:/GeoSpatialData/BiogeographicalRegions/Norway_VegetationZones_Moen/Original/Vector/soner.shp",
               "/data/R/GeoSpatialData/BiogeographicalRegions/Norway_VegetationZones_Moen/Original/Vector/soner.shp")
```

```{r}
soner <- sf::read_sf(path)
```

```{r}
tm_shape(soner)+
  tm_polygons(col = "NAVN")
```

It could make sense to combine some zones, but I will later split the sones by region, and then merge sones that are very uncommon in a spesific region.

#### Regions
Importing a shape file with the regional delineation. 
```{r}
reg <- sf::st_read("data/regions.shp")
#st_crs(reg)
```
